What's in a Domain? Anaylsis of URL Features

### 2022-10-09 paper/paper.tex
Paper update

We design a feature extraction library for URLs that is sensitive to the various sections of URL
structure outlined above. We then apply this library to multiple machine learning tasks using only URLs
as the input variable. We experiment with multiple standard machine learning techniques and evaluate
the impact of the URL features using the SHAP package for feature importance.

@@

These URL feature functions are available in our open source implmentation and python package \emph{url2features}. 
The package is designed such that it can process large files in chunks and will only add the specific sub-groups of
features a user requests.

\subsection{Data}

The data used in this study was colated from a range of sources to represent a varity of 
Internet resource classification problems containing URLs as the primary feature.
The ISCX Malcious URL classification dataset contains malware, spam and phishing URLS 
that need to be discriminated against a set of benign URLS\cite{Mamun2016}.
The world wide web knowledge base (WebKb) 4 Universities data set contains a wide
range of university URLs categorised into multiple topic categories\cite{Craven1998}.
The Syskill \& Webert webpage ratings dataset containing webepages across 4 categories
with a human generated categorisation for determining personal preferences in webpage
content\cite{Pazzani1996}. Finally, we utilise the Kaggle DMOZ dataset that contains 
a large set of topic categorised URLs\cite{dmoz}.

@@

It should be clear from this table that the internet security probelms are much more amenable to URL based 
classification. All three of these problems result in high performing classifiers. In comparison the topic
and user preference based tasks achieve far more modest performance. These problems are hindered by both higher 
cardinality targets, and 2 or the 3 data sets being much smaller. However, the results suggest that in general
URL based topic classification is likely to be difficult.

We take the best performing model for each problem and generate the SHAP values for the records in the test data.
These records are then used to generate the URL Structure Feature Importance Plot shown in Figure \ref{fig:importance}.

@@

We see that for all tasks in these experiments the `path` segment provides the most consistent contribution. The subdomain
and global fetaures appear to be strong for all of the malicious URL detection problems.

>>>>

We design a feature extraction library for URLs that is sensitive to the various sections of URL
structure outlined above. Previous work has presented datasets for specific problems where the features
are provided with a similar structural grouping\cite{Vrban2020}. However, we provide our feature extraction
application designed so that any specific subset of these features can be extracted for a specific task.

We apply this library to multiple machine learning tasks using only URLs as the input variable. We choose
these tasks such that they span very different classes of problems an dataset sizes. 
We experiment with multiple standard machine learning techniques and evaluate
the impact of the URL features using the SHAP package for feature importance. We develop a process
for analysing thess features within the logical structure of the URL.

\subsection{Data}
The data used in this study was colated from a range of sources to represent a varity of
Internet resource classification problems containing URLs as the primary feature.
The ISCX Malcious URL classification dataset contains malware, spam and phishing URLS
that need to be discriminated against a set of benign URLS\cite{Mamun2016}.
The world wide web knowledge base (WebKb) 4 Universities data set contains a wide
range of university URLs categorised into multiple topic categories\cite{Craven1998}.
The Syskill \& Webert webpage ratings dataset containing webepages across 4 categories
with a human generated categorisation for determining personal preferences in webpage
content\cite{Pazzani1996}. Finally, we utilise the Kaggle DMOZ dataset that contains
a large set of topic categorised URLs\cite{dmoz}.

@@

These URL feature functions are available in our open source implmentation and python package \emph{url2features}. 
The package is designed such that it can process large files in chunks and will only add the specific sub-groups of
features a user requests. In addition, the naming convention is designed so that features relating to specific
regions of the URL can be identified and analysed as a group.

@@

@@

The table of results indicates that internet security problems appear much more amenable to URL based 
classification. All three of these problems result in high performing classifiers. In comparison the topic
and user preference based tasks achieve far more modest performance. It is worth noting that these problems 
are hindered by both higher cardinality targets and greater variability in dataset size, 2 of the 3 data 
sets are much smaller than any of the security datasets. However, the performance on the very large DMOZ
dataset strongly suggests that, in general, URL based topic classification is likely to remain a difficult
problem.

We used the results from the best performing classification model for each individual problem and generated
the SHAP values for the records in the test dataset. These records are then used to generate the URL Structure 
Feature Importance Plot shown in Figure \ref{fig:importance}. This plot aggregates the importance of individual
features into their region of origin within the URL structure. Allowing us to visualise the source of signal for
each task and compare features across tasks.

We see that there is substantial variability across all classification tasks in these experiments.
The protocol and parameters sections remain mostly devoid of signal, with the exception of the Spam URL classifcation
task, where we see some value in the URL parameter features. The domain, path and file regions are the predominant sources
of signal across these tasks. Nevertheless, the emphasis changes between them. The subdomain appears particularly
important for the phishing URL identfication, whereas the URL target file is the focus for the WebKb classifciation data.

The internet security problems appear to derive less utility from the global features when compared to the three
topic classification problems, as is indicated by the strength of the grey lines through the middle of the plot 
for each task. The contributions of the domain name itself and the top level domain vary across the tasks and do
not appear consistant across the top broad categories of secutity and topic classification. Suggesting that developers
need to test these features on a task by task basis rather than relying on heuristics.
